\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}

\newcommand{\s}[1]{\mathcal{#1}}

\title{Notes about Differential Geometry}
\author{Jon}
\date{\today}

\begin{document}
\maketitle
\section{What is a manifold?}
\begin{definition}
A topological space $X$ is called \textbf{locally Euclidean} if for every point $x \in X$ there exists a neighborhood $U$ of $x$ such that $U$ is homeomorphic to a subset $U'$ of $\mathbb{R}^n$. We call a pair $(U, \phi)$ where $\phi$ is such a homeomorphism a \textbf{chart}.
\end{definition}

\begin{definition}
A $\mathcal{C}^k$-\textbf{atlas} on a locally Euclidean space $X$ is a set of charts $\{(U_i, \phi_i)\}$ such that $\cup_iU_i = X$ and $\phi_i$ is a homeomorphism from $U_i$ to an open subset $U_i'$ of $\mathbb{R}^n$, and with the property that for any $i$ and $j$ $\phi_i \circ \phi_j^{-1}$ is in $\mathcal{C}^k(U_i \cap U_j)$. 
\end{definition}

\begin{definition}
Two atlases $\mathcal{F}$ and $\mathcal{g}$ are equivalent if $\s{F} \cup \s{G}$ is an atlas.
\end{definition}

\begin{proposition}
Equivalence between atlases is an equivalence relation.
\end{proposition}

\begin{proof}
Reflexivity and Symmetry are obvious. For transitivity, suppose that $\mathcal{F}$, $\mathcal{G}$, and $\mathcal{H}$ are atlases such that $\s{F} \cup \s{G}$ and $\s{G} \cup \s{H}$ are also atlases. Suppose that $(U, h)$ and $(W, \ell)$ are charts in $\s{F}$ and $\s{H}$, respectively. Then for all points $p \in U \cap W$, there exists an atlas $(V_p, k_p)$ in $\mathbb{G}$ such that $p \in V_p$. Then it follows that 
\[(h \circ \ell^{-1}) = (h \circ k_p^{-1}) \circ (k_p \circ \ell^{-1})\] is $\s{C}^k$ differentiable in a neighborhood around $p$ for all $p$ in $U \cap W$, as desired.
\end{proof}

\begin{definition}
Suppose that $\s{A}$ is an atlas. Then a \textbf{differentiable structure} generated by $\s{A}$ is the union of all atlases such that their union with $\s{A}$ is an atlas.
\end{definition}

\begin{definition}
A $\s{C}^k$-differentiable manifold is a pair $(M, \s{F})$, where $M$ is a Hausdorff second countable locally Euclidean topological space, and $\s{F}$ is a differentiable structure.
\end{definition}


\end{document}